Causal Compatibility Inequalities Admitting Quantum Violations in the Triangle Structure

### 2020-06-02 CCIAQVTS/CCIAQVTS.tex

\begin{document}
    \title{Causal Compatibility Inequalities Admitting Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tfraser@perimeterinstitute.ca}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \affiliation{University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1}
    \author{Elie Wolfe}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    % \email{ewolfe@perimeterinstitute.ca}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are incompatible with particular assumption about classical causal structure. Given a causal structure of unknown classicality, the presence of such correlations certifies the non-classical nature of the causal structure in a device independent fashion. In structures where all parties share a common resource, these non-classical correlations are also known as non-local correlations. Any constraint satisfied by all correlations which \textit{are} classically compatible with a given causal structure defines a causal compatibility criterion. Such criteria were recently derived for the Triangle structure (E. Wolfe \textit{et al.}, \href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}) in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution in this work, however, is the derivation of causal compatibility inequalities for the Triangle structure which do admit quantum violation. This provides a robust-to-noise witness of quantum correlations in the Triangle structure. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    % \tableofcontents
    % \setlength{\parskip}{1em}

    \section{Introduction}
    \label{sec:introduction}
    In recent decades, the technical utility of quantum mechanics has become abundantly clear. In the realm of computation, quantum algorithms --- such as Shor's algorithm~\cite{Shor_1997} and numerous others~\cite{Jordan_2016} --- scale exponentially better than their classical counterparts. In the realm of secure communication, quantum protocols --- a popular example being Quantum Key Distribution~\cite{Bennett_2014} --- are able to provide privacy even against hypothetical adversaries with unlimited computational power, a desideratum which classical protocols are unable to fulfill. Throughout history, numerous quantum phenomena which fail to be emulated by classical physics have been identified as resources for solving computational or communicational problems~\cite{Neilsen_Chaung_2011}. Motivated by past successes, a primal objective of modern quantum information theory is to discover new situations wherein quantum mechanics offers an advantage, and to certify that the quantum advantage is genuine.

    From a foundational prospective, the most robust demonstrations of quantum phenomena with no classical emulation have involved Bell inequalities~\cite{Bell_1964,Brunner_2013}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous EPR paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic method for demonstrating the \textit{non-classicality} of a given observation. More recently it has been appreciated that Bell inequalities can understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatibility inequalities, such as Bell and Instrumental inequalities \cite{pearl1995instrumental,bonet2001instrumental,evans2012graphical}, characterize the space of observations that are compatible with a hypothesized causal structure, albeit the characterization offered by practically derivable inequalities is often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell structure, as depicted in \cref{fig:bell_structure}. The fundamental Bell structure involves non-communicating parties making measurements on some hidden shared resource $\la$, where the measurement outcomes ($A, B$) are presumed to be stochastic functions of the local choices of measurement settings ($S_A, S_B$) and the shared resource $\la$. Quantum non-classicality in the Bell structure has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex structures, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular correlation scenario named the Triangle structure (\cref{fig:triangle_structure}).

    The Triangle structure (\cref{fig:triangle_structure}) is a causal structure comprised of $3$ parties labeled $A, B, C$ arranged in a triangular configuration while pair-wise sharing hidden/latent variables $X, Y, Z$. It has been extensively studied previously (see~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, App. E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 4]{Weilenmann_2016},~\cite[Fig. 1]{Inflation}, $\ldots$). An overview of some milestone results is provided in \cref{sec:triangle_structure}. Identifying causal compatibility inequalities for this configuration has been seen as particularly challenging~\cite{Branciard_2012}. Further identifying causal compatibility inequalities of such high resolution that such inequalities can be violated by quantum-accessible distributions has remained out-of-reach for the Triangle structure.

    This work is the first, therefore, to find causal compatibility inequalities for the Triangle structure that are known to be violated by quantum-accessible distributions. This accomplishment was made possible through the combination of two previous developments. First, the insight of \citet{Fritz_2012} regarding the ability to re-interpret the Bell structure as a portion of the Triangle structure. Second, the new framework for solving causal inference problems developed by~\citet{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a validation that the Inflation Technique is efficient and sensitive enough at low orders to offer new insights into quantum non-classicality \cite{Navascues_2017}. Moreover, these new inequalities offer an avenue for recognizing previously unknown forms of non-classicality. The authors attempts to find such novel resources were met with only partial success, suggesting the need for both conceptual refinements and future exploration.

    The first half of the manuscript, namely \Cref{sec:causal_compatibility,sec:bell_structure,sec:triangle_structure,sec:fritz_distribution}, is entirely a review of previous works. \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:bell_structure} offers a summary of the popular Bell structure and associated inequalities. \Cref{sec:triangle_structure} discusses the Triangle structure and provides an overview of existing research; identifying its stark differences from the Bell structure and motivating why the Triangle structure is worth studying. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first conceived of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. In the same work~\cite{Fritz_2012}, the Fritz distribution was proven to be non-classical without the use of inequalities.

    The second half of the manuscript, namely \Cref{sec:preliminary_research,sec:found_inequalities,sec:optimizations,sec:fritz_problem_revisit}, presents the main contributions made by this research project. Specifically, in \cref{sec:found_inequalities} we improve upon the results of \citet{Fritz_2012} by offering a direct proof of the incompatibility of the Fritz distribution using inequalities generated by the Inflation Technique~\cite{Inflation}. A sample of such inequalities are presented in \cref{sec:found_inequalities}, specifically \cref{eq:ww_ineq,eq:web_inequality,eq:symmetric_web_inequality}. Aside from confirming the utility of the Inflation Technique, this paper explores the importance of having derived these inequalities. First, an inequality-based proof has the advantage of being robust to experimental noise. In \cref{sec:violations_noise} the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities. Second, we numerically optimize our derived inequalities over quantum-accessible distributions (using qubits) in an effort to find the maximum violations achievable by quantum theory. The culmination of these analyses naturally prompts a discussion, found in \Cref{sec:fritz_problem_revisit}, regarding the fundamental problem of recognizing and classifying novel non-classicality in the Triangle structure. \Cref{sec:conclusions} concludes.

@@

    If a given distribution $\prob[\nodes]$ defined over \textit{all} of the nodes $\nodes$ of $\graph$ satisfies \cref{eq:compatibility_as_factorization}, then $\prob[\nodes]$ is said to be \term{compatible with $\graph$}. The conditional distributions in \cref{eq:compatibility_as_factorization} (i.e. $\{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes\}$) are referred to as a set of \term{causal parameters for $\graph$}. If a distribution $\prob[\nodes]$ cannot be factorized according to \cref{eq:compatibility_as_factorization}, $\prob[\nodes]$ is said to be \term{incompatible} with $\graph$. When one is specified with a joint distribution $\prob[\nodes]$ defined over all nodes of a causal structure $\graph$, it is possible to completely determine whether or not $\prob[\nodes]$ is compatible with $\graph$ by computing the causal parameters induced by $\prob[\nodes]$ and checking the equality of \cref{eq:compatibility_as_factorization}. A challenge, however, is presented when one is supplied with a partial observation, i.e. a joint distribution $\prob[\nodes_{O}]$ where $\nodes_{O} \subset \nodes$ is some subset of variables referred to as the \term{observable nodes} $\nodes_O$. In such cases, $\prob[\nodes_{O}]$ does not induce a unique set of causal parameters for $\graph$ and \cref{eq:compatibility_as_factorization} can not be verified by direction calculation. Instead, compatibility between $\graph$ and $\prob[\nodes_{O}]$ depends on the existence or non-existence of a set of causal parameters for $\graph$ such that $\prob[\nodes_{O}] = \prod_{n \in \nodes_{O}} \prob[n \mid \Pa[\graph]{n}]$. The complementary, unobservable nodes are termed \term{latent nodes} $\nodes_L = \nodes \setminus \nodes_{O}$, and should be understood as hidden random variables that are unknowable either by some fundamental process or cannot be measured due to other limitations.

@@

    From the perspective of identifying quantum non-classicality, a causal structure $\graph$ adopts the role of a classical hypothesis. Therefore, non-classicality becomes synonymous with incompatibility: if a distribution $\prob[\nodes_{O}]$ is incompatible, then it is non-classical. Henceforth, we will use these two terms interchangeably. From a resource standpoint, if the non-classicality of $\prob[\nodes_{O}]$ with $\graph$ can be witnessed by an inequality $I$, but nevertheless $\prob[\nodes_{O}]$ can be implemented using quantum states and measurements while otherwise respecting the causal relations of $\graph$, then the causal compatibility inequality $I$ represents a task or game where quantum resources out-perform classical resources relative to $\graph$.

    \section{Bell Structure}
    \label{sec:bell_structure}
    This section aims to define the Bell causal structure and to review some of the traditional witnesses used to assess the classicality (or lack thereof) of distributions relative to it. The purpose of this section is to equip readers with the pertinent background and also to draw comparisons between the advancements made toward understanding non-classicality of the Bell structure versus analogous results obtained for Triangle structure later in this work.

@@

    Any observed distribution $\prob[AB|S_AS_B]$ which fails to be explained by the classical causal hypothesis of the Bell structure as defined by \cref{eq:bellcompat} is appropriately termed non-classical. Often, distributions incompatible with the Bell structure are referred to as non-local because the Bell structure markedly lacks causal influence from one party to another. In particular,~\citet{Bell_1964} demonstrated that there exists distributions that are non-classical yet are attainable via local measurements on a shared quantum resource.

    Let us now contrast the classical definition of compatibility with its quantum generalization. Quantum correlation arise from a quantumified version of the causal structure, in which latent variables are replaced with quantum systems. Concretely, a distribution is accessible with quantum resources and measurements for the quantumified Bell structure if and only if there exists a bipartite quantum state $\rho_{AB}$ of arbitrary dimension, as well as Hilbert-space-localized measurement sets $M_{A|S_A}, M_{B|S_B}$\footnote{Note that a measurement set $M_{A} = \bc{M_{A}^{1}, M_{A}^{2}, \ldots, M_{A}^{k}}$ services a shorthand notation in the following sense, $\prob[A] = \Tr\bs{\rho M_{A}} \implies \prob[A]\br{a} = \Tr\bs{\rho M_{A}^{a}}$.} with indices conditional upon the measurement settings, such that
    \begin{align}\label{eq:bellQcompat}
    \prob[AB|S_AS_B] \text{ is a quantum realization of \cref{fig:bell_structure}} \quad\iff\quad\prob[AB|S_AS_B] = \Tr\bs{\rho_{AB}M_{A|S_A}\otimes M_{B|S_B}}\,.
    \end{align}
    The non-classicality of quantum distributions for the Bell structure can be demonstrated through the use of Bell inequalities which constrain the correlations between binary variables $A$ and $B$ for classically compatible distributions. A notable example is the CHSH inequality (\cref{eq:CHSH_ineq_orig})~\cite{CHSH_Original}.
    \[ \ba{AB | S_{A} = 0, S_{B} = 0} + \ba{AB | S_{A} = 0, S_{B} = 1} + \ba{AB | S_{A} = 1, S_{B} = 0} - \ba{AB | S_{A} = 1, S_{B} = 1} \leq 2 \eq \label[ineq]{eq:CHSH_ineq_orig} \]
    Non-trivial Bell inequalities such as the CHSH inequality are capable of witnessing the non-classical nature of quantum distributions; the inequalities presented in \cref{sec:found_inequalities} are also of this type. Equality constraints, as previously mentioned, never have that sort of high-resolution discernment sensitivity.

@@

    The Triangle structure serves as an excellent test case for furthering our understanding of quantum non-classicality in network causal structures. It maintains superficial simplicity (only three observable variables) while introducing many challenging features not found in the study of the Bell structure.
    For example, the spaces of both classical and quantum distributions on the Triangle structure are non-convex \cite{Fritz_2012,Inflation}, unlike for the Bell structure. The convexity of the Bell structure's distributions is arguably responsible for the wealth of knowledge about it, including its complete characterization of classicality \cite{Brunner_2013}.
    Importantly, \citet{Fritz_2012} explicitly demonstrated the existence of (at least) one incompatible-but-quantum distribution for the Triangle structure, so its known to possess quantum non-classicality. It seems reasonable to assume that quantum non-classicality in the Triangle structure should be translatable into a computational advantage for certain computational circuits~\cite{Terhal_2002}; novel instances of non-classicality are expected to correspond to novel information-theoretical quantum advantages. A fundamental limitation of Fritz's proof of non-classicality, however, is that it does not involve causal compatibility inequalities, and hence does not advance our repertoire of inequality constraints for the Triangle structure.
    Some inequality constraints for the Triangle structure have been derived in previous works. For example, \citet{Steudel_2010} derived an inequality distinguishing the distributions compatible with the Triangle structure from those compatible with structures in which all the observable variables share a common latent ancestor. \citet{Henson_2014} derived a family of entropic inequalities for the Triangle structure, which was the expanded somewhat by \citet{Weilenmann_2016}. Recently, \citet{Inflation} derived a variety of new, especially sensitive, polynomial causal compatibility inequalities for the Triangle structure. In particular, the inequalities of \cite{Inflation} expose a previously unclassified (as assessed by all formerly known constraints) distribution called the $w$-distribution\footnote{Although the $w$-distribution is non-classical, it is also non-quantum. The non-quantum nature of the $w$-distribution has been demonstrated by Miguel Navascués and Elie Wolfe in private correspondence.} as incompatible with the Triangle structures.
    Remarkably, \emph{none} of the existing causal compatibility inequalities for the Triangle structure were known to admit violation by any quantum distribution. That is to say, it was unknown if any inequality known for the Triangle structure might be useful for distinguishing quantum distributions from their classical counterparts. \Cref{sec:preliminary_research} reports our attempt to utilize these aforementioned inequalities to search for incompatible distributions that are also quantum-accessible. The subsequent failure of these approaches effectively motivates the remainder of this work.
    In summary, the Triangle structure is a desirable case study because it is known to admit non-classical distributions using quantum resources, but yet no inequality heretofore could separate its classical distributions from its quantum distributions. This failure represents a gap in our understanding of quantum non-classicality and prompts the discovery of \textit{new} inequalities.

@@

    The analysis of non-classicality changes, however, when $S_{A}, S_{B}$ are not freely chosen by the observers but rather by a process outside of the individual party's control. Relaxing the assumption of measurement independence opens up a possible loophole, namely the possibility that the auxiliary outcomes $S_{A}, S_{B}$ of Alice and Bob might be manipulated via dependence on their shared latent variable $\lambda$. This loophole is closed by having the third party in the Triangle structure, Charlie, \emph{also} report the latent variables $S_{A}, S_{B}$ as a multivariate outcome. In this manner, the perfect correlation of $C$'s record of $S_{A}, S_{B}$ with the records of $S_{A}$ reported by $A$ and of $S_{B}$ reported by $B$ testifies to the independence of $S_{A}, S_{B}$ from $\lambda$. Consequently, any distribution over $A, B, S_{A}, S_{B}$ that is incompatible with Bell structure is also incompatible with the Triangle structure provided that $C$ is \textit{perfectly correlated} with $S_{A}, S_{B}$~\cite{Fritz_2012}.

    The exemplifying quantum distribution corresponding to a recasting of a non-classical Bell structure distribution into the Triangle structure is the \term{Fritz distribution}~\cite{Fritz_2012}. In the Fritz distribution, denoted $\prob[\fritz]$, each of the variables $A,B,C$ is taken to have $4$ possible outcomes $\bc{0,1,2,3}$. Explicitly, $\prob[\fritz]$ can be written as:

@@

    We emphasize that each of these causal compatibility inequalities independently provides a positive resolution of Fritz's problem. The inequalities are derived without making use of Bell's theorem; rather, they follow from the Inflation Technique's broader perspective on non-classicality as a special case of causal inference. Afterwards, Section~\ref{sec:fritz_problem_revisit} returns to the topic of Fritz's problem and the status of our resolution.

>>>>

\begin{document}
    \title{Causal Compatibility Inequalities Admitting Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tfraser@perimeterinstitute.ca}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \affiliation{University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1}
    \author{Elie Wolfe}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    % \email{ewolfe@perimeterinstitute.ca}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are incompatible with a particular assumption about classical causal structure. Given a causal structure of unknown classicality, the presence of such correlations certifies the nonclassical nature of the causal structure in a device-independent fashion. In structures where all parties share a common resource, these nonclassical correlations are also known as non-local correlations. Any constraint satisfied by all correlations which are classically compatible with a given causal structure defines a causal compatibility criterion. Such criteria were recently derived for the Triangle structure (E. Wolfe \textit{et al.}, \href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}) in the form of polynomial inequalities, begging the question of whether any of those inequalities admit violation by quantum correlations. Numerical investigation suggests that they do not, and we further conjecture that the set of correlations admitted by the classical Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution in this work, however, is the derivation of causal compatibility inequalities for the Triangle structure which do admit quantum violation. This provides a robust-to-noise witness of quantum correlations in the Triangle structure. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    % \tableofcontents
    % \setlength{\parskip}{1em}

    \section{Introduction}
    \label{sec:introduction}
    In recent decades, the technical utility of quantum mechanics has become abundantly clear. In the realm of computation, quantum algorithms, such as Shor's algorithm~\cite{Shor_1997} and numerous others~\cite{Jordan_2016}, scale exponentially better than their classical counterparts. In the realm of secure communication, quantum protocols, a popular example being quantum key distribution~\cite{Bennett_2014}, are able to provide privacy even against hypothetical adversaries with unlimited computational power, a desideratum which classical protocols are unable to fulfill. Throughout history, numerous quantum phenomena which fail to be emulated by classical physics have been identified as resources for solving computational or communication problems~\cite{Neilsen_Chaung_2011}. Motivated by past successes, a primal objective of modern quantum information theory is to discover new situations wherein quantum mechanics offers an advantage, and to certify that the quantum advantage is genuine.

    From a foundational prospective, the most robust demonstrations of quantum phenomena with no classical emulation have involved Bell inequalities~\cite{Bell_1964,Brunner_2013}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous Einstein-Podolsky-Rosen paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic method for demonstrating the nonclassicality of a given observation. More recently it has been appreciated that Bell inequalities can be understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatibility inequalities, such as Bell and Instrumental inequalities \cite{pearl1995instrumental,bonet2001instrumental,evans2012graphical}, characterize the space of observations that are compatible with a hypothesized causal structure, albeit the characterization offered by practically derivable inequalities is often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell structure, as depicted in \cref{fig:bell_structure}. The fundamental Bell structure involves noncommunicating parties making measurements on some hidden shared resource $\la$, where the measurement outcomes ($A$ and $B$) are presumed to be stochastic functions of the local choices of measurement settings ($S_A$ and $S_B$) and the shared resource $\la$. Quantum nonclassicality in the Bell structure has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex structures, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular correlation scenario named the Triangle structure (\cref{fig:triangle_structure}).

    The Triangle structure (\cref{fig:triangle_structure}) is a causal structure comprised of three parties labeled $A, B,$ and $C$ arranged in a triangular configuration while pairwise sharing hidden (latent) variables $X, Y,$ and $Z$. It has been extensively studied previously (see, e.g.,~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, App. E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 4]{Weilenmann_2016}, and~\cite[Fig. 1]{Inflation}). An overview of some milestone results is provided in \cref{sec:triangle_structure}. Identifying causal compatibility inequalities for this configuration has been seen as particularly challenging~\cite{Branciard_2012}. Further identifying causal compatibility inequalities of such high resolution that such inequalities can be violated by quantum-accessible distributions has remained out-of-reach for the Triangle structure.

    This work is the first, therefore, to find causal compatibility inequalities for the Triangle structure that are known to be violated by quantum-accessible distributions. This accomplishment was made possible through the combination of two previous developments. First, the insight of \citet{Fritz_2012} regarding the ability to re-interpret the Bell structure as a portion of the Triangle structure. Second, the new framework for solving causal inference problems developed by~\citet{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a validation that the Inflation Technique is efficient and sensitive enough at low orders to offer new insights into quantum nonclassicality \cite{Navascues_2017}. Moreover, these new inequalities offer an avenue for recognizing previously unknown forms of nonclassicality. The authors attempts to find such novel resources were met with only partial success, suggesting the need for both conceptual refinements and future exploration.

    The first half of the manuscript, namely \Cref{sec:causal_compatibility,sec:bell_structure,sec:triangle_structure,sec:fritz_distribution}, is entirely a review of previous works. \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:bell_structure} offers a summary of the popular Bell structure and associated inequalities. \Cref{sec:triangle_structure} discusses the Triangle structure and provides an overview of existing research; identifying its stark differences from the Bell structure and motivating why the Triangle structure is worth studying. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first conceived of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. In the same work~\cite{Fritz_2012}, the Fritz distribution was proven to be nonclassical without the use of inequalities.

    The second half of the manuscript, namely \Cref{sec:preliminary_research,sec:found_inequalities,sec:optimizations,sec:fritz_problem_revisit}, presents the main contributions made by this research project. Specifically, in \cref{sec:found_inequalities} we improve upon the results of \citet{Fritz_2012} by offering a direct proof of the incompatibility of the Fritz distribution using inequalities generated by the Inflation Technique~\cite{Inflation}. A sample of such inequalities are presented in \cref{sec:found_inequalities}, specifically \cref{eq:ww_ineq,eq:web_inequality,eq:symmetric_web_inequality}. Aside from confirming the utility of the Inflation Technique, this paper explores the importance of having derived these inequalities. First, an inequality-based proof has the advantage of being robust to experimental noise. In \cref{sec:violations_noise} the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities. Second, we numerically optimize our derived inequalities over quantum-accessible distributions (using qubits) in an effort to find the maximum violations achievable by quantum theory. The culmination of these analyses naturally prompts a discussion, found in \Cref{sec:fritz_problem_revisit}, regarding the fundamental problem of recognizing and classifying novel nonclassicality in the Triangle structure. \Cref{sec:conclusions} concludes.

@@

    If a given distribution $\prob[\nodes]$ defined over \textit{all} of the nodes $\nodes$ of $\graph$ satisfies \cref{eq:compatibility_as_factorization}, then $\prob[\nodes]$ is said to be \term{compatible with $\graph$}. The conditional distributions in \cref{eq:compatibility_as_factorization} (i.e. $\{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes\}$) are referred to as a set of \term{causal parameters for $\graph$}. If a distribution $\prob[\nodes]$ cannot be factorized according to \cref{eq:compatibility_as_factorization}, $\prob[\nodes]$ is said to be \term{incompatible} with $\graph$. When one is specified with a joint distribution $\prob[\nodes]$ defined over all nodes of a causal structure $\graph$, it is possible to completely determine whether or not $\prob[\nodes]$ is compatible with $\graph$ by computing the causal parameters induced by $\prob[\nodes]$ and checking the equality of \cref{eq:compatibility_as_factorization}. A challenge, however, is presented when one is supplied with a partial observation, i.e. a joint distribution $\prob[\nodes_{O}]$ where $\nodes_{O} \subset \nodes$ is some subset of variables referred to as the \term{observable nodes} $\nodes_O$. In such cases, $\prob[\nodes_{O}]$ does not induce a unique set of causal parameters for $\graph$ and \cref{eq:compatibility_as_factorization} can not be verified by direction calculation. Instead, compatibility between $\graph$ and $\prob[\nodes_{O}]$ depends on the existence or non-existence of a set of causal parameters for $\graph$ such that $\prob[\nodes_{O}] = \sum_{n \not \in \nodes_{O}} \prob[\nodes]$ where $\prob[\nodes]$ is again given by \cref{eq:compatibility_as_factorization}. The complementary, unobservable nodes are termed \term{latent nodes} $\nodes_L = \nodes \setminus \nodes_{O}$, and should be understood as hidden random variables that are unknowable either by some fundamental process or cannot be measured due to other limitations.

@@

    From the perspective of identifying quantum nonclassicality, a causal structure $\graph$ adopts the role of a classical hypothesis. Therefore, nonclassicality becomes synonymous with incompatibility: if a distribution $\prob[\nodes_{O}]$ is incompatible, then it is nonclassical. Henceforth, we will use these two terms interchangeably. From a resource standpoint, if the nonclassicality of $\prob[\nodes_{O}]$ with $\graph$ can be witnessed by an inequality $I$, but nevertheless $\prob[\nodes_{O}]$ can be implemented using quantum states and measurements while otherwise respecting the causal relations of $\graph$, then the causal compatibility inequality $I$ represents a task or game where quantum resources out-perform classical resources relative to $\graph$.

    \section{Bell Structure}
    \label{sec:bell_structure}
    This section aims to define the Bell causal structure and to review some of the traditional witnesses used to assess the classicality (or lack thereof) of distributions relative to it. The purpose of this section is to equip readers with the pertinent background and also to draw comparisons between the advancements made toward understanding nonclassicality of the Bell structure versus analogous results obtained for Triangle structure later in this work.

@@

    Any observed distribution $\prob[AB|S_AS_B]$ which fails to be explained by the classical causal hypothesis of the Bell structure as defined by \cref{eq:bellcompat} is appropriately termed nonclassical. Often, distributions incompatible with the Bell structure are referred to as non-local because the Bell structure markedly lacks causal influence from one party to another. In particular,~\citet{Bell_1964} demonstrated that there exists distributions that are nonclassical yet are attainable via local measurements on a shared quantum resource.

    Let us now contrast the classical definition of compatibility with its quantum generalization. Quantum correlation arise from a quantumified version of the causal structure, in which latent variables are replaced with quantum systems. Concretely, a distribution is accessible with quantum resources and measurements for the quantumified Bell structure if and only if there exists a bipartite quantum state $\rho_{AB}$ of arbitrary dimension, as well as Hilbert-space-localized measurement sets $M_{A|S_A}, M_{B|S_B}$\footnote{Note that a measurement set $M_{A} = \bc{M_{A}^{1}, M_{A}^{2}, \ldots, M_{A}^{k}}$ services a shorthand notation in the following sense, $\prob[A] = \Tr\bs{\rho M_{A}} \implies \prob[A]\br{a} = \Tr\bs{\rho M_{A}^{a}}$.} with indices conditional upon the measurement settings, such that
    \begin{align}\label{eq:bellQcompat}
    \prob[AB|S_AS_B] \text{ is a quantum realization of \cref{fig:bell_structure}} \quad\iff\quad\prob[AB|S_AS_B] = \Tr\bs{\rho_{AB}M_{A|S_A}\otimes M_{B|S_B}}\,.
    \end{align}
    The nonclassicality of quantum distributions for the Bell structure can be demonstrated through the use of Bell inequalities which constrain the correlations between binary variables $A$ and $B$ for classically compatible distributions. A notable example is the CHSH inequality (\cref{eq:CHSH_ineq_orig})~\cite{CHSH_Original}.
    \[ \ba{AB | S_{A} = 0, S_{B} = 0} + \ba{AB | S_{A} = 0, S_{B} = 1} + \ba{AB | S_{A} = 1, S_{B} = 0} - \ba{AB | S_{A} = 1, S_{B} = 1} \leq 2 \eq \label[ineq]{eq:CHSH_ineq_orig} \]
    Non-trivial Bell inequalities such as the CHSH inequality are capable of witnessing the nonclassical nature of quantum distributions; the inequalities presented in \cref{sec:found_inequalities} are also of this type. Equality constraints, as previously mentioned, never have that sort of high-resolution discernment sensitivity.

@@

    The Triangle structure serves as an excellent test case for furthering our understanding of quantum nonclassicality in network causal structures. It maintains superficial simplicity (only three observable variables) while introducing many challenging features not found in the study of the Bell structure.
    For example, the spaces of both classical and quantum distributions on the Triangle structure are non-convex \cite{Fritz_2012,Inflation}, unlike for the Bell structure. The convexity of the Bell structure's distributions is arguably responsible for the wealth of knowledge about it, including its complete characterization of classicality \cite{Brunner_2013}.
    Importantly, \citet{Fritz_2012} explicitly demonstrated the existence of (at least) one incompatible-but-quantum distribution for the Triangle structure, so its known to possess quantum nonclassicality. It seems reasonable to assume that quantum nonclassicality in the Triangle structure should be translatable into a computational advantage for certain computational circuits~\cite{Terhal_2002}; novel instances of nonclassicality are expected to correspond to novel information-theoretical quantum advantages. A fundamental limitation of Fritz's proof of nonclassicality, however, is that it does not involve causal compatibility inequalities, and hence does not advance our repertoire of inequality constraints for the Triangle structure.
    Some inequality constraints for the Triangle structure have been derived in previous works. For example, \citet{Steudel_2010} derived an inequality distinguishing the distributions compatible with the Triangle structure from those compatible with structures in which all the observable variables share a common latent ancestor. \citet{Henson_2014} derived a family of entropic inequalities for the Triangle structure, which was the expanded somewhat by \citet{Weilenmann_2016}. Recently, \citet{Inflation} derived a variety of new, especially sensitive, polynomial causal compatibility inequalities for the Triangle structure. In particular, the inequalities of \cite{Inflation} expose a previously unclassified (as assessed by all formerly known constraints) distribution called the $w$-distribution\footnote{Although the $w$-distribution is nonclassical, it is also non-quantum. The non-quantum nature of the $w$-distribution has been demonstrated by Miguel Navascués and Elie Wolfe in private correspondence.} as incompatible with the Triangle structures.
    Remarkably, \emph{none} of the existing causal compatibility inequalities for the Triangle structure were known to admit violation by any quantum distribution. That is to say, it was unknown if any inequality known for the Triangle structure might be useful for distinguishing quantum distributions from their classical counterparts. \Cref{sec:preliminary_research} reports our attempt to utilize these aforementioned inequalities to search for incompatible distributions that are also quantum-accessible. The subsequent failure of these approaches effectively motivates the remainder of this work.
    In summary, the Triangle structure is a desirable case study because it is known to admit nonclassical distributions using quantum resources, but yet no inequality heretofore could separate its classical distributions from its quantum distributions. This failure represents a gap in our understanding of quantum nonclassicality and prompts the discovery of \textit{new} inequalities.

@@

    The analysis of nonclassicality changes, however, when $S_{A}, S_{B}$ are not freely chosen by the observers but rather by a process outside of the individual party's control. Relaxing the assumption of measurement independence opens up a possible loophole, namely the possibility that the auxiliary outcomes $S_{A}, S_{B}$ of Alice and Bob might be manipulated via dependence on their shared latent variable $\lambda$. This loophole is closed by having the third party in the Triangle structure, Charlie, \emph{also} report the latent variables $S_{A}, S_{B}$ as a multivariate outcome. In this manner, the perfect correlation of $C$'s record of $S_{A}, S_{B}$ with the records of $S_{A}$ reported by $A$ and of $S_{B}$ reported by $B$ testifies to the independence of $S_{A}, S_{B}$ from $\lambda$. Consequently, any distribution over $A, B, S_{A}, S_{B}$ that is incompatible with Bell structure is also incompatible with the Triangle structure provided that $C$ is \textit{perfectly correlated} with $S_{A}, S_{B}$~\cite{Fritz_2012}.

    The exemplifying quantum distribution corresponding to a recasting of a nonclassical Bell structure distribution into the Triangle structure is the \term{Fritz distribution}~\cite{Fritz_2012}. In the Fritz distribution, denoted $\prob[\fritz]$, each of the variables $A,B,C$ is taken to have $4$ possible outcomes $\bc{0,1,2,3}$. Explicitly, $\prob[\fritz]$ can be written as:

@@

    We emphasize that each of these causal compatibility inequalities independently provides a positive resolution of Fritz's problem. The inequalities are derived without making use of Bell's theorem; rather, they follow from the Inflation Technique's broader perspective on nonclassicality as a special case of causal inference. Afterwards, Section~\ref{sec:fritz_problem_revisit} returns to the topic of Fritz's problem and the status of our resolution.
